---
title: "Computing model-free area under the curve (AUC)"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Computing model-free area under the curve (AUC)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(tempodisco)
```

"Area under the curve" (AUC) is often used as an empirical measure of discounting ((Myerson et al., 2001)[https://doi.org/10.1901/jeab.2001.76-235]). To compute this measure, we first fit a model using the `'model-free'` discount function (which identifies indifference points separately at each delay), and then call the `AUC` function. For example:

```{r}
# Get data containing indifference points
data("adj_amt_sim")
df <- adj_amt_indiffs(adj_amt_sim)
# Fit model
mod <- td_ipm(df, discount_function = 'model-free')
print(AUC(mod))
```

To compute the AUC only between certain limits, you can use the `min_del` and `max_del` arguments:

```{r}
print(AUC(mod, min_del = 1, max_del = 100))
```

For the AUC computed this way, the later indifference points tend to have an outsize influence on the overall measure. To address this, [Borges et al. (2016)](https://doi.org/10.1002/jeab.219) suggest transforming the delays to a log or ordinal scale. After this transformation, we can re-fit the model and re-compute the AUC. For example, for the log transformation:

```{r}
df$untransformed_del <- df$del # Save a copy
df$del <- log(df$untransformed_del)
mod <- td_ipm(df, discount_function = 'model-free')
print(AUC(mod))
```

Similarly, for the ordinal transformation:

```{r}
df$del <- as.numeric(factor(df$untransformed_del, ordered = T))
mod <- td_ipm(df, discount_function = 'model-free')
print(AUC(mod))
```

Note that any transformation of the delays must be performed _prior_ to fitting the model. This is because, for the model-free discount function, interpolations between indifference points are _linear_ in the original scale of the data. This is evident in the curved interpolations in the plot below:

```{r}
df$del <- df$untransformed_del
mod <- td_ipm(df, discount_function = 'model-free')
plot(mod, log = 'x', verbose = F)
```
